---
output:
  md_document:
    variant: markdown_github
---

```{r, echo = FALSE}
knitr::opts_chunk$set(
    collapse = TRUE,
    comment = "#>",
    fig.path = "inst/README-"
)
```

# jdify

jdify is an R package implementing classifiers based on the joint density of 
the predictors and the class variable. Several methods for joint density estimation
can be used.

To install, open R and type
``` r
devtools::install_github("tnagler/jdify")
```

### Functionality

#### Classification modeling

The core function in this package is `jdify()` which builds a classification
model for a given data set. It estimates the joint density of the predictors
and the class variable and derives conditional class probabilities from it.
```{r, include = FALSE}
library(jdify)
```

```{r}
dat <- data.frame(
    cl = as.factor(rbinom(10, 1, 0.5)),
    x1 = rnorm(10),
    x2 = ordered(rbinom(10, 5, 0.3))
)
model <- jdify(cl ~ x1 + x2, data = dat, jd_method = "cctools")
probs <- predict(model, dat, what = "probs")  # conditional probabilities
```


`jdify()` can handle discrete predictors. They have to be declared as `ordered` or 
`factor` (for unordered categorical variables). All other variables are treated
as continuous.

#### Methods for joint density estimation

You can choose from three built-in methods for: `"cctools"` (default), 
`"kdevine"`, `"np"`. The method name indicates the package that is used for
joint density estimation.  

You can also create custom functions for density estimation by `jd_method()`. 
The following is another implementation of the method `"kdevine"`.

```{r}
my_fit <- function(x, ...)
   kdevine::kdevine(x, ...)
my_eval <- function(object, newdata, ...)
   kdevine::dkdevine(newdata, object)
my_method <- jd_method(fit_fun = my_fit, eval_fun = my_eval, cc = TRUE)
model <- jdify(cl ~ x1 + x2, data = dat, jd_method = my_method)
```

The option `cc = TRUE` indicates that the method does not naturally handle 
discrete data. In this case, `jdify` automatically invokes the continuous
convolution trick (see, Nagler, 2017).

#### Cross validation and performance assessment

`cv_jdify()` is a convenience function that does k-fold cross validation for 
you. It splits the data, fits joint density models and evaluates the conditional
class probabilities on the hold-out samples.

```{r}
cv <- cv_jdify(cl ~ x1 + x2, data = dat, folds = 3)
cv$cv_probs
```
The function `assess_clsfyr()` allows to calculate several performance measures
from the conditional class probabilities. Its first argument is the probability
of the class, the second is a class indicator. 
```{r}
assess_clsfyr(cv$cv_probs[, 1], dat[, 1] == 0, measure = c("ACC", "F1"))
```


### References 

Nagler, T. (2017). *Nonparametric density estimation with discrete variables.*
Unpublished manuscript.

